---
title: "Exercise 1"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # this will make your code visible along with your plots
knitr::opts_chunk$set(message = FALSE) # removes messages from your final output
knitr::opts_chunk$set(warning = FALSE) # removes warnings from your final output
library(tidyverse)
library(here)
```

# Objectives

+  View data 
+  Create histogram of oyster lengths
+  Work in layers
+  Facets

## About the data

This data was collected by the Guana Tolomato Matanzas National Estuarine Research Reserve as part of an oyster monitoring pilot project led by Head Biologist Pamela Marcum. A report with this information can be found [here](https://gtmnerr.org/wp-content/uploads/2018/10/2014-2016-GTM-oyster-monitoring.pdf).

We are going to use this data to visualize the distribution of oyster lengths within different regions of the GTMNERR as well as within harvest and non-harvest zones.

# A

Read in oysters data by running the chunk below and examine the variables in your dataset.

```{r, include = FALSE}
load(here('data', 'oysters.RData'))


```

1.  What are the variables in `oysters`?

[Add your answer here]

2.  What class are the variables?

[Add your answer here]

3.  What are the unique regions in which this data was collected? (*Hint: check out the `unique()` function for help*)

[Add your answer here]

4.  When were these data collected? Explore the date variable.

[Add your answer here]

# B

Next, create a histogram of oyster lengths for the whole dataset using the height measurements.

```{r}


```

Note the warning message about picking better `binwidth`. Let's change the binwidth to 5.

```{r}

```

# C

Fishery-sized oysters ("harvestable") must be three inches in length (76 mm). Let's add a vertical line to the plot using `geom_vline` at 76mm for fishery-size.

```{r}


```

Excellent! Now, the upper limit of oyster spat (newly settled oysters) is 25mm. Let's add a line on there for that, too, but make it a "dashed" linetype.

```{r}


```

# D

Next, let's look at how these oyster lengths look within and outside of harvest zones by filling the bars by the `Harvest` variable.

```{r}


```

# E

Using facet_wrap, separate the harvest information into two separate plots

```{r}


```

***

What are your observations of this data?

[Add your answer here]
