---
title: "Visualize Data: The Basics"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: yeti
    highlight: tango
    df_print: paged
editor_options:
  chunk_output_type: console
---

<!-- This file by Shannon Dunnigan was adapted from the original work at https://github.com/rstudio/master-the-tidyverse by RStudio. -->

# Introduction to Data Visualization

Data visualization is the first step in data exploration and is also a great place to start off when learning `R` because it can be very satisfying when learning to code - "*I want to make a plot, and I did it!*! We will be using the `ggplot2` package. This is one of the earliest packages in the `tidyverse` and it implements the "grammar of graphics", which is a logical system for describing and building graphs.

# The Basics Part 1

+  Mappings and aesthetics
+  Facets
+  Geoms
+  Stat
+  Position
+  Global vs Local settings

"r setup" is an optional label for a chunk, but it should always run once before anything else. Notice that in this setup we are loading the `tidyverse` and `here` packages.

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(here)
```

The `wq` data frame contains monthly averages of water quality data (e.g., temperature, salinity, dissolved oxygen, etc.) from six National Estuarine Research Reserves in the contiguous United States. *Notice the use of the `here` package in the code below.*

```{r loaddata}
wq <- read.csv(here::here('data', 'monthly_wq.csv'))

```

**Tip**: Always be sure to check your data. You can use `head()`, `tail()`, or `str()` from the base package in `R`. Or one of our favorite quick an easy ways is `glimpse()`, which is a great all in one snapshot of the data.

**Quick Quiz**

What relationship do you expect to see between temperature (temp) and dissolved oxygen (do_mgl)?

```{r your-turn-1}
ggplot(data = wq) +
	geom_point(mapping = aes(x = temp, y = do_mgl))

```

There are two main parts to this scatterplot:

1.    Initialize the figure with `ggplot()`. This creates a coordinate system for which you can then add subsequent layers. The first argument of this function is the dataset to use in the graph, so `ggplot(data = <DATA>)` creates an empty graph. (*go on, you can try just plotting that*)

2.    You add layers with `geom_` functions. Each `geom_` function takes a mapping argument. This is how the variables are defined and mapped to visual properties. The mapping argument is always paired with `aes()` and the x and y arguments of `aes()` specify which variables to map to the x and y axes by looking at the `<DATA>` argument in the initial code.

This gives us the start of a nice template for future work with `ggplot2`:

```{r ggplot2-template, eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

***

## What are Mappings?

**Aesthetics** are the visual properties of the objects in your plot (e.g., color, size, shape, etc.). **Mappings** describe how the aethetics should *relate to variables* in the data.

For example, you could have the points in a scatterplot be colored by their associated "state". In visual space, we see the colors of each point, but in data space each of these colors represents a data value from the "state" variable.

```{r aes-example1, echo = FALSE}
ggplot(wq) +
  geom_point(aes(x = temp, y = do_mgl, color = state), size = 3)
```

Each geom comes with their own suite of aesthetics that it recognizes. You can use the `?geom_<insert>` function to explore the associated aesthetics for each geom.

You can also set the aesthetic properties of a geom manually, but be careful when doing so. Anything within the `aes()` call maps an aesthetic to a variable. This means that `R` will be looking for that variable and the data associated within it in order to create the plot. Outside of the `aes()` call, maps an aesthetic to a value. As an example, run the two lines of code below.

```{r aes-example}
ggplot(wq) +
	geom_point(aes(x = temp, y = do_mgl, color = "blue"))


ggplot(wq) +
	geom_point(aes(x = temp, y = do_mgl), color = "blue")
```

What did you notice about the two plots?

## Facets

What if you don't want to add additional variables as aesthetics? If they are categorical variables, you can use **facets**! Facets create subplots that each display one subset of the data. There are two types of facets in `ggplot2`: `facet_wrap()` and `facet_grid()`.

If you only have a single variable, it's best to use `facet_wrap()`. If you have a combination of two variables, use `facet_grid()`.

```{r facet-example}
ggplot(data = wq) +
  geom_point(mapping = aes(x = temp, y = do_mgl)) +
  facet_wrap(~ coast, nrow = 1)

ggplot(data = wq) +
  geom_point(mapping = aes(x = temp, y = do_mgl)) +
  facet_grid(. ~ state)
```

What did the period do in the `facet_grid()` example?

## Geoms

### Basics

The geometric object ('geom') is the visual object used to represent the data. Two plots may share all the same data, but in using different geometric objects, the visual representation of the data will appear a little different. Take the following plots for example:

```{r geom-example, echo = FALSE, message = FALSE}
ggplot(wq) +
  geom_point(aes(x = temp, y = do_mgl))

ggplot(wq) +
  geom_smooth(aes(x = temp, y = do_mgl))

```

Notice how both plots have the same x and y variables as well as the data - they just have different visual objects (one is a scatterplot, the other is a smoothed plot).

Every geom requires a mapping argument and they each come with their own suite of acceptable aesthetics. You can find the list of available aesthetics for each geom within their help file or even by using the `ggplot2` cheatsheet from RStudio.

#### Example 1

Replace this scatterplot with one that draws boxplots.

```{r}
# scatterplot
ggplot(data = wq) +
  geom_point(mapping = aes(x = coast, y = temp))

# boxplot
ggplot(data = wq) +
  geom_boxplot(mapping = aes(x = coast, y = temp))

# geom_boxplot also has aesthetics you can map
ggplot(data = wq) +
  geom_boxplot(mapping = aes(x = coast, y = temp, color = state))

ggplot(data = wq) +
  geom_boxplot(mapping = aes(x = coast, y = temp, fill = state))

ggplot(data = wq) +
  geom_boxplot(mapping = aes(x = coast, y = temp, linetype = coast))

```

Let's check out another dataset called `diamonds`. This data set contains prices and other attributes of diamonds and is provided by the `ggplot2` package - so we do not need to load it like we did the water quality dataset.

#### Example 2

Load the `diamonds` dataset and examine it. *ahem* Remember the `glimpse()` function? This data set contains prices and other attributes of diamonds and is provided by the `ggplot2` package.

Create a bar chart of `cut` colored by `cut`. Use the cheatsheet.

```{r}
diamonds

# always good to check your data
glimpse(diamonds)

ggplot(diamonds) +
  geom_bar(aes(x = cut, color = cut))

```

Hmmm, was this what you had imagined? No, not what I was wanting either. When using the "color" attribute with bar charts, it does not "fill" the bars, it simply provides an outline of color to them. If you want the bars to be "filled" with color, you'll want to use the "fill" aesthetic that `geom_bar` recognizes.

```{r bar-example}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```

That's better...except...wait! Check out the y-axis; "count" is not a variable in `diamonds`!
So, what just happened?

Many graphs plot the raw values of your dataset while others calculate new values to plot:
-  Bar charts, histograms, and frequency polygons bin data and plot bin counts (# of points that fall into each bin)
-  Smoothers, `geom_smooth()`, fit a model to your data and plot the predictions from the model
-  Boxplots compute a summary of the distribution and display it in a formatted box

In short, `geom_bar()` begins with the `diamonds` dataset. It then transforms the data with the "count" stat, which returns a data set of cut values and counts. It then uses the transformed data to build the plot where "cut" is mapped to the x axis and count to the y.

## Stat

Short for ???statistical transformation??? - the algorithm that is used to create new values for a graph. It is also one of the five components of each layer!

```{r, eval = FALSE}
ggplot(data = <DATA>) +
	<GEOM_FUNCTION>(mapping = aes(<MAPPINGS>),
			      stat = <STAT>,
			      position = <POSITION>)
```

## Position

Position is how overlapping objects are arranged in the plot. Instead of just filling our bars with information from 'cut' let's use 'clarity'.

```{r position-example}
# geom_bar defaults to "stacking" the variable infromation. You can look at the help information or the cheatsheet to see what the default stat is for each geom function.
ggplot(data = diamonds) +
	geom_bar(mapping = aes (x = cut, fill = clarity))

# how was fill different? - works like stacking, but each set of stacked bars are the same height. Much better for looking at proportions
ggplot(data = diamonds) +
	geom_bar(mapping = aes (x = cut, fill = clarity),
		    position = "fill")

# how was dodge different? - Places overlapping objects side by side, making it easier to compare individual values
ggplot(data = diamonds) +
	geom_bar(mapping = aes (x = cut, fill = clarity),
		    position = "dodge")

```

## Global versus Local

What if you have several layers you want to apply in a plot and they all use the same basic aesthetics (e.g. x, y, etc)? It would be quite time consuming to write all of those out in each geom_ function argument. Therefore, you can apply some of them globally to all layers by including them in the `ggplot()` code. Then, mappings applied in specific geom_ functions will add or override the global mappings for that layer only.

```{r global-local-example, message = FALSE}
ggplot(data = wq, mapping = aes(x = temp, y = do_mgl)) +
	geom_point() +
	geom_smooth()

ggplot(data = wq, mapping = aes(x = temp, y = do_mgl)) +
	geom_point(mapping = aes(color = coast)) +
	geom_smooth()

```

Notice how only color is applied to the points in the figure, not anywhere else. That is because color was only an aesthetic that was applied in the geom_point layer.


# Part 1 Recap:

To make a graph:

1.    Pick your *data* set
2.    Choose the best *geom* to display the cases
3.    *Map* aesthetic properties to variables.
4.    **Not required**, but defaults are supplied by the specified geom that you can then manually change.

```{r, eval = FALSE}
ggplot(data = <DATA>) +
	<GEOM FUNCTION>(mapping = aes(<MAPPINGS>),
				   stat = <STAT>, # not required, but sensible default applied.
				   position = <POSITION>) # not required, but sensible default applied.
```

# The Basics Part 2

What else can you do?

+  Scales
+  Coordinate Systems
+  Themes
+  Saving your plots

## Scales

*Scales* control the way your data is mapped to your geom and are required by every aesthetic. Again, `ggplot2` selects sensible defaults, but you can provide further adjustments.

```{r scales-example}
ggplot(wq, aes(x = temp, y = do_mgl)) +
  geom_point(aes(shape = coast, color = state), size = 3)

ggplot(wq, aes(x = temp, y = do_mgl)) +
  geom_point(aes(shape = coast, color = state), size = 3) +
  scale_shape_discrete(name = "US Coast") + # renamed scale for the shape aesthetic, applies to legend
  scale_color_discrete(name = "US State") + # renamed scale for the color aesthetic, applies to legend
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10)) + # adjusted the axis breaks in the y axis scale
  scale_x_continuous(breaks = c(5, 10, 15, 20, 25, 30, 35)) # adjusted the axis breaks in the x axis scale

# manual
ggplot(wq, aes(x = temp, y = do_mgl)) +
  geom_point(aes(color = coast), size = 3) +
  scale_color_manual(labels = c('Atlantic', 'Gulf', 'Pacific'),
                     values = c('#47A49D', '#D9582E', '#B0BC40')) 

# New update to changing legend names
ggplot(wq, aes(x = temp, y = do_mgl)) +
  geom_point(aes(shape = coast, color = state), size = 3) +
  scale_y_continuous(breaks = c(2, 4, 6, 8, 10)) +
  scale_x_continuous(breaks = c(5, 10, 15, 20, 25, 30, 35)) +
  labs(color = "US State", # add aesthetic into the labs() argument!
       shape = "US Coast")

```

## Coordinate Systems

Coordinate systems combine two position aesthetics to produce a 2D position on a plot. In a typical scatterplot these may be `x` and `y`, but in a polar coordinate system these are likely angle and radius. 

The Cartesian coordinate system is the most familiar and common of the coordinate systems. Setting limits on the coordinate system will zoom in the plot without changing the underlying data like setting limits in a scale will.

```{r coord-linear-example}
ggplot(data = wq) +
  geom_point(aes(x = temp, y = do_mgl, color = state), size = 3) +
  coord_cartesian(xlim = c(0,15),
                  ylim = c(7, 11))
```

There are two types of coordinate systems: linear and non-linear. Non-linear coordinate systems may change the shape of the geoms. 

Let's check out some linear options using the `diamonds` data from earlier.

```{r coord-linear-example2}
# classic and default Cartesian coordinate system with x and y limits
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))

# let's flip the x and y axes
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut)) +
  coord_flip()

```

We'll convert the `diamonds` data into a polar coordinate system.

```{r coord-polar-example}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = '', fill = cut),
           position = "fill") +
  coord_polar(theta = "y") +
  theme_void()

```

## Titles and captions

These are really quite easy - just add a `labs()` layer!

```{r titles-example}
ggplot(data = wq) +
 geom_boxplot(mapping = aes(x = coast, y = temp, color = state)) +
 labs(x = "Coast",
      y = "Temperature (C)",
      title = "Water Quality Data",
      subtitle = "Pacific Coast Reserves are colder",
      caption = "Data from <www.nerrsdata.org>")
```

## Themes

These are the non-data elements of your plot and do not affect how the data is:
-   rendered by geoms, or
-   transformed by scales.

Themes help you make the plot aesthetically pleasing or match an existing style guide (fonts, ticks, panel strips, backgrounds, etc). Notice how the data, and the aesthetics associated with the data like color, do not change between the following four plots. Only the non-data elements (e.g., axes, gridlines, background colors, etc.) are changing. These are the theme elements.

```{r themes-example, echo = FALSE}
a <- ggplot(wq) +
  geom_point(aes(x = temp, y = do_mgl, color = state), size = 3) +
  theme_bw()
b <- ggplot(wq) +
  geom_point(aes(x = temp, y = do_mgl, color = state), size = 3) +
  theme_classic()
c <- ggplot(wq) +
  geom_point(aes(x = temp, y = do_mgl, color = state), size = 3) +
  theme_dark()
d <- ggplot(wq) +
  geom_point(aes(x = temp, y = do_mgl, color = state), size = 3) +
  theme_minimal()

cowplot::plot_grid(a, b, c, d, labels = "AUTO")
rm(a, b, c, d)
```

The structure of the theme system goes like this:

Three components:

1.   `theme()` function
2.   Theme elements (e.g., `plot.title` element which controls the appearance of the plot title)
3.   Element function (of which there are four)

      -   `element_text()`
      -   `element_line()`
      -   `element_rectangle()`
      -   `element_blank()`

```{r themes-example-template, eval = FALSE}
p + theme(plot.title = element_text(size = 18),
	     axis.y.title = element_text(size = 12),
	     axis.x.title = element_text(size = 12),
	     axis.text = element_text(vjust = 0.9),
	     panel.background = element_rect(fill = ???white???))
```

Explore all the different elements by going to the help page with `?theme`.

## Saving your plots

*What does `getwd()` return?* 

```{r}
getwd() # returns your working directory (i.e., where R is looking for and saving your files.)
```

### The Working Directory

`R` associates itself with a folder (i.e., directory) on your computer.
-   This is the ???working directory???
-   R will look for files here
-   R will save files here

The files pane of RStudio IDE will display your working directory

```{r, eval = FALSE}
# ggsave() will save the last plot.
   ggsave("my-first-plot.png")
   ggsave("my-first-plot.tiff")

# You can further specify size:
   ggsave("my-first-plot.png", height = 6, width = 8,
           units = "cm")

# Or even resolution:
	ggsave("my-first-plot.png", dpi = 300)
```

# Take aways

You can use this code template to make thousands of graphs with `ggplot2`.

```{r, eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>),
                  stat = <STAT>,
                  position = <POSITION>) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION> +
  <SCALE_FUNCTION> +
  <THEME_FUNCTION>
```

