---
title: "Visualize Data: Beyond the Basics"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: yeti
    highlight: tango
    df_print: paged
editor_options:
  chunk_output_type: console
---

<!-- This file by Shannon Dunnigan was adapted from the original work at https://github.com/rstudio/master-the-tidyverse by RStudio. -->

# The Basics Recap

+  Mappings and aesthetics
+  Facets
+  Geoms
+  Stat
+  Position
+  Global vs Local settings
+  Scales
+  Coordinate Systems
+  Themes
+  Saving your plots

To make a graph:

1.    Pick your *data* set
2.    Choose the best *geom* to display the cases
3.    *Map* aesthetic properties to variables.
4.    **Not required**, but defaults are supplied by the specified geom that you can then manually change.

You can use this code template to make thousands of graphs with `ggplot2`.

```{r, eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>),
                  stat = <STAT>,
                  position = <POSITION>) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION> +
  <SCALE_FUNCTION> +
  <THEME_FUNCTION>
```

# Beyond the Basics

+  Conditional terms in ggplot
+  Multiplots with `patchwork`
+  Interactive plots with `plotly`
+  Theme customization with `ggThemeAssist`
+  Adding plot statistics with `ggpubr` *maybe covered, but time dependent.


"r setup" is an optional label for a chunk, but it should always run once before anything else. Notice that in this setup we are loading the `tidyverse` and `here` packages.

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(here)
library(patchwork)
library(plotly)
library(ggThemeAssist)
library(ggpubr)
```

The `guana_nut` data frame contains monthly extracted chlorophyll a data collected within the Guana Lake and River system from 2017-2021 in Ponte Vedra Beach, Florida. This data is provided by the Guana Tolomato Matanzas National Estuarine Research Reserve and is currently maintained by Olivia Roorbach. *Notice the use of the `here` package in the code below.*

```{r loaddata}
load(file = here('data', 'guana_nut.RData'))

```

## Conditional Aesthetics 

We can also set conditions inside of aesthetic properties which is of particular help when telling stories through data. Let's look at some chlorophyll a data. Also, I want to point out that you can use filtering functions on a dataset inside the ggplot code to consolidate your coding effort!

```{r cond-1}
ggplot(data = filter(guana_nut, site_acronym == "GR1"), 
       aes(x = sample_date, y = result)) +
	geom_line() +
  geom_point() +
  labs(x = "Date",
       y = "Chlorophyll a (\U00B5g/L)")

```

Wow, there are some high values in there. The Florida State threshold for chlorophyll a in Class II Estuary waters is 6.6ug/L. Let's add a threshold line and pull out all of the values that exceed that threshold. 

```{r cond-2}
ggplot(data = filter(guana_nut, site_acronym == "GR1"), 
       aes(x = sample_date, y = result)) +
	geom_line() +
  geom_point(aes(color = result > 6.6), size = 2) +
  geom_hline(yintercept = 6.6, linetype = "dashed") +
  labs(x = "Date",
       y = "Chlorophyll a (\U00B5g/L)")

```

You can also change the aesthetics from color to shape, etc. 

These can also further be defined in the scales and provided with specified labels and values.

```{r cond-3}
ggplot(data = filter(guana_nut, site_acronym == "GR1"), 
       aes(x = sample_date, y = result)) +
	geom_line() +
  geom_point(aes(color = result > 6.6), size = 2) +
  geom_hline(yintercept = 6.6, linetype = "dashed") +
  scale_color_manual(name = "",
                     labels = c("< 6.6 \U00B5g/L", "> 6.6 \U00B5g/L"),
                     values = c("black", "red")) +
  labs(x = "Date",
       y = "Chlorophyll a (\U00B5g/L)")

```

## Multiplots with `patchwork`

The `patchwork` package makes creating multiplots as easy as basic math...because it's like math! Remember PEMDAS from grade school? Anything inside of parentheses will be done first.

```{r}
GR1 <- ggplot(data = filter(guana_nut, site_acronym == "GR1"), 
       aes(x = sample_date, y = result)) +
	geom_line() +
  geom_point(aes(color = result > 6.6), size = 2) +
  geom_hline(yintercept = 6.6, linetype = "dashed") +
  scale_color_manual(name = "",
                     labels = c("< 6.6 \U00B5g/L", "> 6.6 \U00B5g/L"),
                     values = c("black", "red")) +
  labs(x = "Date",
       y = "Chlorophyll a (\U00B5g/L)")

GR <- ggplot(data = filter(guana_nut, site_acronym == "GR"), 
       aes(x = sample_date, y = result)) +
	geom_line() +
  geom_point(aes(color = result > 6.6), size = 2) +
  geom_hline(yintercept = 6.6, linetype = "dashed") +
  scale_color_manual(name = "",
                     labels = c("< 6.6 \U00B5g/L", "> 6.6 \U00B5g/L"),
                     values = c("black", "red")) +
  labs(x = "Date",
       y = "Chlorophyll a (\U00B5g/L)")

River_sites <- ggplot(data = filter(guana_nut, wbid == "River"), 
       aes(x = sample_date, y = result, color = site_acronym, group = site_acronym)) +
	geom_line() +
  geom_point(size = 2) +
  labs(x = "Date",
       y = "Chlorophyll a (\U00B5g/L)",
       color = "Site")

# notice the single sites figures are on the top and next to one another and the full series is underneath the two
(GR1 + GR) / River_sites
```

## Interactive plots with `plotly`

Use the `plotly` package to turn any of your ggplots into interactive plots with the `ggplotly()` function. Simply wrap the plot inside the function and it will output into the "Viewer" tab of your RStudio!

## Theme customization with `ggThemeAssist`

When you install `ggThemeAssist` and open the package with `library()`, it adds into your RStudio as an "Addin", which you can navigate to under the main navigation bar on the far right, typically under the "Help" tab. 

